---
title: "Logistic Regression Analysis of Topical and Rhetorical Framing Features"
author: "Qi Yu"
date: "2023-02-20"
output:
  html_document: default
  pdf_document: default
---

```{r message=FALSE}
library(tidyverse)
library(broom)
library(dplyr)
library(tibble)
library(ggplot2)
library(ggsci)
library(visreg)
library(jtools)
library(lme4)
library(corrplot)
library(Hmisc)
```

# 1. Read in feature statistics
```{r}
df <- read.csv("/Users/qiyu/PycharmProjects/topical-and-rhetorical-framing/linguistic_annotation_service/results/feature_statistics.tsv", sep="\t", header=TRUE)
save_root = "results/"
```


# 2. Preprocessing 

Combine the statistics of all discourse connective types to one feature ```connective```: 
```{r}
df$connective <- df$causal + df$consecutive + df$conditional + df$concessive + df$adversative
```

Remove columns that will not be used for logistic regression:
```{r}
df <- select(df, -c(id, text, text_cleaned,
                    causal, consecutive, conditional, concessive, adversative
                    )) 
```


Rearrange the column orders, so that features belonging to the same type and dimension (see Section 4 of the original paper) are groupped together: 
```{r}
col_order <- c("source", 
               "arousal", # dimension "arousal"
               "adv_iter_cont", "scalar_particle",# dimension "presupposition"
               "common_ground", "resigned_accept", "weak_commit", # dimension "modal particles"
               "exclamation", "question", # dimension "sentence type"
               "connective", # dimension "information structure"
               "economy", "identity", "legal", "morality", "policy", "politics", "public_opinion", "security", "welfare" # topical framing features
               )

df <- df[, col_order]
```

# 3. Check correlation of the features

```{r results = 'hide'}
res <-rcorr(as.matrix(df[, 2:ncol(df)]), type="spearman")
sort(c(round(as.matrix(res$r), 3)))
```
```{r}
label_col = rep(c("#08519C", "black"), times = c(9, 9))
corrplot(res$r, type="upper", method="number", tl.col = label_col, tl.srt = 45, number.cex= 0.5, tl.cex = 1, cl.cex = 1, mar = c(0, 0, 0, 0))
```

As ```economy``` and ```welfare``` show a relatively strong correlation whereas the vocabulary of ```economy``` has a broader coverage, we discard the feature ```welfare```:
```{r}
df <- select(df, -c(welfare))
```

# 4. Logistic regression

To quantify which features are most distinctive in each newspaper source (BILD, FAZ and SZ), 
for each source we fit a binary logistic regression model using the source as
response variable (e.g., is_BILD = 0 vs.1) and all features as predictors.

Firstly, we create the binary response variable: 
```{r}
df <- df %>% 
  mutate(is_bild = if_else(source == 0, "yes", "no"), .after = "source") %>% 
  mutate(is_faz = if_else(source == 1, "yes", "no"), .after = is_bild) %>% 
  mutate(is_sz = if_else(source == 2, "yes", "no"), .after = is_faz) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(source = factor(source))
```

Standardize all predictors by removing the mean and scaling to unit variance:
```{r}
x <- df[, 5: ncol(df)] %>%
  mutate_if(is.numeric, scale)
x <- as.data.frame(x)
```

Fit logistic regression model for each newspaper source. 

Results of all significant features are written out to the folder /results.
```{r}
sources <- c("bild", "faz", "sz")

for (s in grep("is_bild", colnames(df)) : grep("is_sz", colnames(df))) {
  y <- df[, s]
  df_logistic <- cbind(y, x)

  model <- glm(y ~ ., data = df_logistic, family = 'binomial')
  print(summary(model))

  result <- tidy(model) %>%
    mutate_if(is.numeric, round, digits=5)

  result <- result[order(result$statistic, decreasing=TRUE), ]
  write.csv(result[result$p.value < 0.05, ], paste(save_root, "logistic_regression_", sources[s-1], ".csv", sep = ""), row.names=FALSE)
}
```

# 5. Plot logistic regression results: 

Read in the result files from Section 4, and group the features into two types (```Topical```, ```Rhetorical```) for plotting:
```{r}
df_bild <- read.csv("./results/logistic_regression_bild.csv") %>% 
  mutate(source = "BILD") %>% 
  mutate(feature_type = if_else(term %in% c("economy", "identity", "legal", "morality", "policy", "politics", "public_opinion", "security", "welfare"), "Topical", "Rhetorical")) %>% 
  filter(term!="(Intercept)")

df_faz <- read.csv("./results/logistic_regression_faz.csv") %>% 
   mutate(source = "FAZ") %>% 
  mutate(feature_type = if_else(term %in% c("economy", "identity", "legal", "morality", "policy", "politics", "public_opinion", "security", "welfare"), "Topical", "Rhetorical")) %>% 
  filter(term!="(Intercept)")

df_sz <- read.csv("./results/logistic_regression_sz.csv") %>% 
   mutate(source = "SZ") %>% 
  mutate(feature_type = if_else(term %in% c("economy", "identity", "legal", "morality", "policy", "politics", "public_opinion", "security", "welfare"), "Topical", "Rhetorical")) %>% 
  filter(term!="(Intercept)")
```

For each newspaper, plot the z-value of each feature. Note that only significant features are included here (see Section 4):
```{r}
ggplot(df_bild, aes(x = reorder(term, statistic), y = statistic, fill = feature_type)) + 
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  ylab("Z-Score") + 
  ggtitle("BILD") + 
  scale_fill_jco() +
  theme_grey() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size = 21), 
        legend.position = "none")
```

```{r}
ggplot(df_faz, aes(x = reorder(term, statistic), y = statistic, fill = feature_type)) + 
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  ylab("Z-Score") + 
  ggtitle("FAZ") + 
  scale_fill_jco() +
  theme_grey() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size = 21),
        legend.position = "none")
```

```{r}
ggplot(df_sz, aes(x = reorder(term, statistic), y = statistic, fill = feature_type)) + 
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  ylab("Z-Score") + 
  ggtitle("SZ") + 
  scale_fill_jco() +
  theme_grey() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size = 21)) + 
  guides(fill=guide_legend(title="Feature Type"))
```




